Structural analysis

### 2024-09-20 workflow/notebooks/draft.Rmd
switched to notung

tax <- read_delim(params_dataset$taxons_file, show_col_types=F)
taxmap <- read_delim(glue(outdir, "/db/taxidmap"), 
                     delim = "\t", col_names = c("label", "Tax_ID"),
                     show_col_types = FALSE) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  dplyr::select(label, Proteome_ID, mnemo)

@@

sptree <- read.tree(params_dataset$species_tree_labels)
translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
# sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]

@@

# read input table
lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))

@@

table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
                    delim = "\t", col_names = table_columns) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
         Clade = factor(Clade, levels = lvls_tax))

lddt_df <- read_delim(paste0(params$structure_dir, "/", tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
           col_names = c("target", "mean_lddt"), show_col_types = FALSE)

@@

```{r cath}
CATH_df <- read_delim(list.files(params$cath_dir, full.names = TRUE),
                      delim = "\t", id = "proteome", show_col_types = FALSE) %>%
  # filter(!is.na(Gene3D)) %>% 
  mutate(proteome = gsub("_cath.tsv", "", basename(proteome)),
         n_3d_domains = str_count(Gene3D, ";"),
         n_Pfam_domains = str_count(Pfam, ";")) %>%
  # filter(n_domains==1) %>% 
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
    select(-proteome, Entry)

nms <- read_delim("test/cath/cath-names.txt", 
                  comment = "#", delim = "    ",
                  col_names = c("Gene3D", "abbr", "name"), 
                  show_col_types = FALSE) %>% 
  select(-abbr) %>% 
  mutate(name=gsub("^:", "", name))

nms_pfam <- read_delim("test/cath/Pfam-A.clans.tsv", 
           col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
           show_col_types = FALSE) %>% 
  select(-name, -clan, -boh)

@@

@@

```{r fig2}
evals <- c(Inf, 1e-2, 1e-3, 1e-5)
min_covs <- c(0, 30, 80)
eval_p <- ggplot(df, aes(-log10(evalue_Blast), -log10(evalue_Foldseek))) + 
  stat_bin_2d(breaks=seq(0,20,0.4)) + 
  scale_fill_gradientn(colors=wespal) +
  geom_hline(yintercept = -log10(evals), linetype=2) + 
  geom_vline(xintercept = -log10(evals), linetype=2) + 
  coord_fixed(expand = 0) +
  labs(x="-log10(Evalue Blast)",y="-log10(Evalue Foldseek)") +
  scale_x_continuous(limits = c(0, 20)) +
  scale_y_continuous(limits = c(0, 20)) +
  theme(legend.position = "none", panel.grid = element_blank()) 
eval_p <- add_marginal(mutate(df, 
                              evalue_Blast=-log10(evalue_Blast), 
                              evalue_Foldseek=-log10(evalue_Foldseek)), 
                       eval_p, "evalue_Blast", "evalue_Foldseek", "singleton", palette_singleton_p)

pident_p <- ggplot(df, aes(pident_Blast, pident_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  scale_fill_gradientn(colors=wespal) +
  # scale_fill_viridis_c(option = "C") +
  coord_fixed(expand = 0) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="Identity Blast",y="Identity Foldseek") +
  theme(legend.position = "none", panel.grid = element_blank())
pident_p <- add_marginal(df, pident_p, "pident_Blast", "pident_Foldseek", "singleton", palette_singleton_p)

qcov_p <- ggplot(df, aes(qcov_Blast, qcov_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  geom_hline(yintercept = min_covs, linetype=2) + 
  geom_vline(xintercept = min_covs, linetype=2) + 
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="Query cov. Blast",y="Query cov. Foldseek") +
  scale_fill_gradientn(colors=wespal) +
  coord_fixed(expand = 0) +
  theme(legend.position = "none", panel.grid = element_blank()) 
qcov_p <- add_marginal(df, qcov_p, "qcov_Blast", "qcov_Foldseek", "singleton", palette_singleton_p)

@@

df_multi <- df_filtered %>% 
    # head(n = 100000) %>%
  left_join(CATH_df, by=c("query"="target")) %>% 
  filter(!is.na(Pfam), !is.na(Gene3D)) %>% 
  left_join(CATH_df, by="target") %>% 
  rowwise() %>% 
  mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
         Pfam.x=strsplit(Pfam.x, ";"),
         Gene3D.y=strsplit(Gene3D.y, ";"),
         Pfam.y=strsplit(Pfam.y, ";")) %>% 
  mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
         n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
  # filter(n_Pfam_common>0, n_3d_common>0) %>% 
  select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
  mutate(Jaccard_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
         Jaccard_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
  select(query, target, singleton, contains("Jaccard"), contains("common"))

CATH_plot <- df_multi %>% 
    pivot_longer(contains("Jaccard"), values_to = "Jaccard") %>% 
    mutate(name=gsub("Jaccard_", "", name),
           name=gsub("3d", "CATH", name),
           Jaccard=cut_width(Jaccard, .1, boundary = 0)) %>% 
  filter(!is.na(Jaccard)) %>% 
  group_by(name, singleton, Jaccard) %>% 
  summarise(n=n()) %>% 
  mutate(freq=n/sum(n)) %>%
  ggplot(aes(Jaccard, freq, fill=singleton)) + 
  geom_bar(stat = "identity",position = "dodge") + 
  facet_grid(name~.) +
  labs(y="% in group") +
  scale_y_continuous(expand = expansion(0.01,0)) + 
  scale_fill_manual(values=palette_singleton_p) +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill=NA, color="black"),
        axis.line = element_blank(),
        axis.text.x = element_text(size=8))

@@

@@

rf_text <- group_by(rf_df, model.x, model.y) %>% 
  filter(as.integer(model.x)<=as.integer(model.y)) %>% 
  summarise(value=median(RF, na.rm=T))

@@

# Reconciliation ranger+TCS scores
reco <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_DTL.tsv"), 
                   show_col_types = FALSE)

scores <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_scores.tsv"), 
                     show_col_types = FALSE) %>% 
  left_join(reco) %>% 
  mutate(model=factor(model, levels=models),
         scores, n_events = (dups+losses)/n_tips)
```

```{r fig3}
my_comparisons <- list( c("LG", "3Di"), c("3Di", "GTR"), c("LG", "GTR") )
plot_bs <- df_bs %>% 
  mutate(targets=target) %>% 
  # group_by(gene, target, model) %>% 
  # summarise(mn_bs=mean(support)) %>% 
  ggplot(aes(model, support, fill=targets)) +
  # geom_violin(position = position_dodge(width = 1)) +
  geom_boxplot(outlier.size = .1) +
  # ggdist::stat_pointinterval(position = position_dodge(width = 1), color="black") +
  scale_fill_manual(values = palettes_method) +
  theme_classic() +
  labs(y="Boostrap", x="") 
  # theme(legend.position = "bottom") 
  # ggpubr::stat_compare_means(aes(group = model),
  #                            comparisons = my_comparisons,
  #                            label = "p.format")

rf_plot <- rf_df %>% 
  ggplot(aes(RF)) +
  geom_tile(aes(x=0.5, y=0.5, fill=value), data = rf_text) +
  geom_text(aes(x=0.5, y=.5, label=round(value, 2)), color="white", data = rf_text) +
  geom_rug(linewidth=.1, data= . %>% filter(as.integer(model.x)>as.integer(model.y))) +
  geom_density(aes(y = after_stat(scaled)), data= . %>% filter(as.integer(model.x)>as.integer(model.y))) + 
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Median RF") + 
  scale_x_continuous(breaks = c(0,.25,.5, .75,1), labels = function(x) sprintf("%.2g", x)) +
  coord_cartesian(xlim = c(0,1), expand=0) +
  # facet_grid(method~method2, switch = "both") +
  facet_grid(model.x~model.y) +
  # ggh4x::facet_nested(targets+model~targets2+model2, nest_line = element_line(linetype = 1)) +
  labs(x="RF Distance") +
  theme_classic() +
  theme(legend.position = "bottom", 
        panel.background = element_rect(fill='transparent'),
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip.text = element_text(size=6),
        axis.text.x = element_text(size=6),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-20,0,0,0))

comps <- NULL
idx <- 1
for (el in df_trees$target %>% unique()){
  for (el2 in df_trees$target %>% unique()){
    if (el>el2){
    comps[[idx]] <- c(el, el2)
    idx <- idx+1
    }
  }
}

ttest <- scores %>%
  filter(model!="FTPY", id %in% id_tmp) %>%
  group_by(model) %>% 
  rstatix::t_test(n_events~targets) %>% 
  rstatix::add_xy_position(x="model", dodge = 0.8)

bxp <- scores %>% 
  filter(model!="FTPY", id %in% id_tmp) %>%
  ggpubr::ggboxplot(x = "model", y = "n_events", fill = "targets",
                    palette = palettes_method)

bxp +
ggpubr::stat_pvalue_manual(ttest, label = "p")

ggplot(scores,aes(n_tips, dups+losses, color=targets)) + 
  geom_point(size=.1) +
  geom_smooth(method = "lm", alpha=.2) +
  facet_wrap(~model, nrow = 1,scales = "free") +
  # coord_cartesian(xlim = c(0,50), ylim = c(0,250)) +
  scale_color_manual(values = palettes_method)

plot_DL <- scores %>% 
  filter(model!="FTPY") %>% 
  ggplot(aes(model, n_events, fill=targets)) +
  geom_boxplot(outlier.size = .1) +
  labs(x="", y="(D+L)/# Tips") +
  # ggthemes::scale_fill_solarized() +
  scale_fill_manual(values = palettes_method)
  # theme(legend.position = "bottom")

var_r2t <- scores %>% 
  filter(model!="FTPY", targets=="common") %>% 
  ggplot(aes(color=model, variance_r2t)) + 
  geom_density(size=1, alpha=.5) +
  # geom_histogram(position = "dodge") + 
  # coord_cartesian(expand = 0) +
  scale_x_continuous(limits = c(0,.2)) + 
  # facet_grid(~targets) + 
  labs(x="Variance Root-to-tip") +
  scale_color_manual(values = palettes_model) +
  theme(legend.position = 'bottom',
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,0,0))

secondcolumn <- (plot_DL + plot_bs + 
    plot_layout(guides = "collect", widths = c(5,3)) &
    theme(legend.position = 'bottom',
          legend.margin=margin(0,0,0,0),
          legend.box.margin=margin(-20,0,0,0))) / var_r2t

fig3 <- (rf_plot | secondcolumn) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
ggsave("draft/figures/fig3.pdf", fig3, width = 11, height = 6)
```

```{r FigS1}
df_red <- select(df_filtered, query, target, singleton)

@@

# for what this should be grouped?
plot_dist <- res %>% 
  filter(query!=target) %>% 
  group_by(query, singleton, model) %>%
  summarise(median_dist=median(nn_dist/max_dist), n=n()) %>% 
  ggplot(aes(median_dist, model,  fill=singleton)) + 
  ggdist::stat_halfeye(alpha=.8, point_colour = NA, .width = 0, height=.6, 
                       position = position_nudge(y=.15)) +
  geom_boxplot(
    width=.2,
    position = position_dodge(),
    # removing outliers
    outlier.color = NA,
    alpha = 1,
    # position= position_nudge(y=-.5)
  ) + 
  scale_x_continuous(expand = c(0,0.01), limits = c(0,1)) +
  theme(panel.border = element_rect(fill = NA)) + 
  scale_fill_manual(values = palette_singleton_p) + 
  theme(legend.position = "bottom")

labeled_sptree <- read.tree("data/sptrees/homo_internal.spTree.nw")
nodes_clades <- fortify(labeled_sptree) %>% 
  # filter(!isTip) %>% 
  mutate(ordered = rank(y)) %>% 
  mutate(label_int = fct_reorder(label, y)) %>%
  select(label_int, ordered)

clades <- read_delim(params_dataset$taxons_file, show_col_types = FALSE)

trees_df <- fortify(apro_trees) 
nodes_df <- trees_df %>% 
  group_by(.id) %>% 
  mutate(ordered = rank(y)) %>% 
  left_join(nodes_clades) %>% 
  separate(.id, c("targets", "model")) %>% 
  filter(!isTip, label!="") %>% 
  mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
  separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), 
           ";", convert = TRUE) %>% 
  mutate(freq=f1+f2+f3)

pp1_plot <- nodes_df %>%
    filter(model!="FTPY") %>% 
    ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
    geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
                   data = . %>% 
                     group_by(model, label_int) %>% 
                     summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
                   position = position_dodge(width=.7), color="grey70") +
    geom_point(aes(shape=targets),
               position = position_dodge(width=.7)) +
  scale_fill_manual(values=palettes_model) +
  scale_color_manual(values = palettes_model_d) +
  scale_shape_manual(values = c(21,22,23,24)) +
  labs(y="Node") +
  guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
         shape = guide_legend(override.aes = list(size=3))) +
  theme(legend.position = "bottom", legend.box="vertical") + 
  theme_bw()

nodes_df %>% 
  ggplot(aes(label_int, f1+f2+f3, fill=model)) + 
  geom_bar(stat="identity", position = "dodge") +
  facet_grid(targets~.) +
  scale_fill_manual(values=palettes_model)

sptree_plot <- ggtree(labeled_sptree) + 
  geom_tiplab() + 
  geom_nodelab(color="firebrick", 
               nudge_x = -.15, hjust = 1, 
               nudge_y = .3, size=2)+
  scale_x_continuous(expand = c(.2,.2))

plot_ntips <- scores %>% 
  filter(model!="FTPY") %>% 
  ggplot(aes(n_tips, dups+losses, color=targets)) + 
  geom_point(size=.2) +
  geom_smooth(method = "lm") +
  facet_grid(~model) + 
  scale_color_manual(values = palettes_method) +
  theme_classic() + 
  theme(panel.border = element_rect(fill=NA),
        legend.position = "bottom")

fig_s1 <- (plot_dist + sptree_plot + pp1_plot) +
  plot_layout(widths = c(1.4, 1, 1.6)) + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 6)

# fig_s1 <- (plot_ntips / ((plot_dist + sptree_plot + pp1_plot) +
#                            plot_layout(widths = c(1.4, 1, 1.6)))) + 
#   plot_layout(heights = c(.75,1)) 
# ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 8)
```

## Answers to Edu's comments

>>>>

tax <- read_delim(glue(outdir, "/meta/", params_dataset$dataset, "_uniprot_genomes.tsv"), 
                  show_col_types=F)
taxmap <- read_delim(glue(outdir, "/db/taxidmap"), 
                     delim = "\t", col_names = c("label", "Tax_ID"),
                     show_col_types = FALSE) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  dplyr::select(label, Proteome_ID, mnemo)

@@

sptree <- read.tree(params_dataset$species_tree)
translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
# sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]

@@

# read input table
# lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))

@@

# table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
#                     delim = "\t", col_names = table_columns) %>% 
#   left_join(tax, by = "Tax_ID") %>% 
#   mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
#          Clade = factor(Clade, levels = lvls_tax))

lddt_df <- read_delim(paste0(params$data_dir, "structures/", 
                             tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
           col_names = c("target", "mean_lddt"), show_col_types = FALSE)

@@

```{r cath}
CATH_df <- read_delim(list.files(glue(params$data_dir, "cath"), full.names = TRUE),
                      delim = "\t", id = "proteome", show_col_types = FALSE) %>%
  # filter(!is.na(Gene3D)) %>% 
  mutate(proteome = gsub("_cath.tsv", "", basename(proteome)),
         n_3d_domains = str_count(Gene3D, ";"),
         n_Pfam_domains = str_count(Pfam, ";")) %>%
  # filter(n_domains==1) %>% 
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
    select(-proteome, Entry)

nms <- read_delim("resources/cath-names.txt", 
                  comment = "#", delim = "    ",
                  col_names = c("Gene3D", "abbr", "name"), 
                  show_col_types = FALSE) %>% 
  select(-abbr) %>% 
  mutate(name=gsub("^:", "", name))

nms_pfam <- read_delim("resources/Pfam-A.clans.tsv", 
           col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
           show_col_types = FALSE) %>% 
  select(-name, -clan, -boh)

@@

## Fig1

```{r fig1}
fig1a <- df %>%
  group_by(singleton) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(prop=n/sum(n), Filtering="Before") %>%
  bind_rows(df_filt %>%
            group_by(singleton) %>%
            summarise(n = n()) %>%
            ungroup() %>%
            mutate(prop=n/sum(n), Filtering="After")) %>%
  mutate(Filtering=factor(Filtering, labels = c("After", "Before"), ordered = T),
         singleton=case_when(singleton=="only Blast"~"Only Bp", 
                             singleton=="only Foldseek" ~ "Only Fs",
                             .default = singleton)) %>%
  ggplot(aes(singleton, n, fill=singleton)) +
  geom_bar(aes(alpha=Filtering), stat = "identity", position = position_dodge2(reverse=TRUE)) +
  geom_text(aes(group=Filtering, label=paste0(round(prop, 2)*100, "%")),
            position = position_dodge2(reverse = TRUE, width = 0.9),
            vjust=1.3, color="white", size=3) +
  scale_fill_manual(values = palette_singleton_s, guide="none") +
  scale_alpha_manual(values = c(1,.6)) +
  scale_y_continuous(expand = c(0,0), labels = scales::label_number(scale = 1e-6)) +
  labs(fill="Class", x="Class", y="# aligned protein pairs (Mln)") +
  theme(legend.position = "bottom", legend.box="vertical")

fig1b <- lddt_df %>% 
  ggplot(aes(mean_lddt)) + 
  geom_density(fill="#BF4375", size=1) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  labs(x="Mean pLDDT", y="") +
  geom_vline(xintercept = 40, linetype=2) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank())

ggsave("draft/figures/fig1_lddt.pdf", fig1b, width = 3, height = 4)
ggsave("draft/figures/fig1_cnts.pdf", fig1a, width = 4, height = 2)
```

@@

```{r fig2}
eval_p <- ggplot(df, aes(-log10(evalue_Blast), -log10(evalue_Foldseek))) + 
  stat_bin_2d(breaks=seq(0,20,0.4)) + 
  scale_fill_gradientn(colors=wespal) +
  ggpubr::stat_cor(label.x.npc = .025) +
  # geom_hline(yintercept = -log10(evals), linetype=2) + 
  # geom_vline(xintercept = -log10(evals), linetype=2) + 
  coord_fixed(expand = 0) +
  labs(x="-log10(Evalue BlastP)",y="-log10(Evalue Foldseek)") +
  scale_x_continuous(limits = c(0, 20)) +
  scale_y_continuous(limits = c(0, 20)) +
  theme(legend.position = "none", panel.grid = element_blank()) 
eval_p <- add_marginal(mutate(df, 
                              evalue_Blast=-log10(evalue_Blast), 
                              evalue_Foldseek=-log10(evalue_Foldseek)), 
                       eval_p, "evalue_Blast", "evalue_Foldseek", "singleton", palette_singleton_p)

pident_p <- ggplot(df, aes(pident_Blast, pident_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  scale_fill_gradientn(colors=wespal) +
  ggpubr::stat_cor(label.x.npc = .025) +
  # scale_fill_viridis_c(option = "C") +
  coord_fixed(expand = 0) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="% Identity BlastP",y="% Identity Foldseek") +
  theme(legend.position = "none", panel.grid = element_blank())
pident_p <- add_marginal(df, pident_p, "pident_Blast", "pident_Foldseek", "singleton", palette_singleton_p)

qcov_p <- ggplot(df, aes(qcov_Blast, qcov_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  ggpubr::stat_cor(label.x.npc = .025) +
  # geom_hline(yintercept = min_covs, linetype=2) + 
  # geom_vline(xintercept = min_covs, linetype=2) + 
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="Query cov. BlastP",y="Query cov. Foldseek") +
  scale_fill_gradientn(colors=wespal) +
  coord_fixed(expand = 0) +
  theme(legend.position = "none", panel.grid = element_blank()) 
qcov_p <- add_marginal(df, qcov_p, "qcov_Blast", "qcov_Foldseek", "singleton", palette_singleton_p)

@@

df_multi <- df_filt %>% 
  select(query, target, singleton) %>% 
    head(n = 1000) %>%
  left_join(CATH_df, by=c("query"="target")) %>% 
  filter(!is.na(Pfam), !is.na(Gene3D)) %>% 
  left_join(CATH_df, by="target") %>% 
  rowwise() %>% 
  mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
         Pfam.x=strsplit(Pfam.x, ";"),
         Gene3D.y=strsplit(Gene3D.y, ";"),
         Pfam.y=strsplit(Pfam.y, ";")) %>% 
  mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
         n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
  # filter(n_Pfam_common>0, n_3d_common>0) %>% 
  select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
  mutate(Jaccard_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
         Jaccard_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
  select(query, target, singleton, contains("Jaccard"), contains("common"))

CATH_plot <- df_multi %>% 
    pivot_longer(contains("Jaccard"), values_to = "Jaccard") %>% 
    mutate(name=gsub("Jaccard_", "", name),
           name=gsub("3d", "CATH", name),
           Jaccard=cut_width(Jaccard, .1, boundary = 0)) %>% 
  filter(!is.na(Jaccard)) %>% 
  group_by(name, singleton, Jaccard) %>% 
  summarise(n=n()) %>% 
  mutate(freq=n/sum(n)) %>%
  ggplot(aes(Jaccard, freq, fill=singleton)) + 
  geom_bar(stat = "identity",position = "dodge") + 
  facet_grid(name~.) +
  labs(y="% in Singleton group", x="Jaccard Index", 
       fill="Target sets") +
  scale_y_continuous(expand = expansion(0.01,0)) + 
  scale_fill_manual(values=palette_singleton_p) +
  theme(legend.position = "top",
        panel.background = element_rect(fill=NA, color="black"),
        axis.line = element_blank(),
        axis.text.x = element_text(size=8))

@@

disco_fls <- list.files(glue(outdir, "/reco/disco"), pattern = "output.nwk", full.names = T)

disco_rf <- NULL
for (file in disco_fls){
  a <- read.tree(file)
  rf <- TreeDist::RobinsonFoulds(sptree, a, normalize = T)
  ntips <- sapply(a, function(x) length(x$tip.label))
  disco_rf <- bind_rows(disco_rf, 
                        tibble(ntips=ntips, rf=rf, bn=rep(basename(file), length(rf))))
}

@@

rf_text <- group_by(rf_df, model.x, model.y) %>% 
  # filter(as.integer(model.x)<=as.integer(model.y)) %>% 
  summarise(value=median(RF, na.rm=T))

@@

n_tips <- sapply(ts, function(x) length(x$tip.label)) %>% 
  enframe() %>%
  separate(name, c("gene", "target", "alphabet", "model")) %>%
  rename("n_tips"="value")

n_sps <- sapply(ts, function(x) 
  length(unique(taxidmap[taxidmap$target %in% x$tip.label, ]$Tax_ID))) %>% 
  enframe() %>%
  separate(name, c("gene", "target", "alphabet", "model")) %>%
  rename("n_sps"="value")

var_df <- sapply(ts, function(x)
  var(adephylo::distRoot(phytools::midpoint_root(x)))) %>% 
  enframe() %>%
  separate(name, c("gene", "target", "alphabet", "model")) %>%
  rename("variance_r2t"="value")

# Reconciliation ranger+TCS scores
reco <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_notung.tsv"), 
                   show_col_types = FALSE, col_names = c("gene", "dups", "losses")) %>% 
  separate(gene, c("gene", "target", "alphabet", "model")) %>% 
  left_join(var_df) %>% 
  left_join(n_tips) %>% 
  left_join(n_sps) %>% 
  mutate(model=factor(model, levels=models),
         n_events = (dups+losses)/n_tips) %>% 
  rename("targets"="target")
```

```{r fig3}

plot_DL <- reco %>% 
  filter(model!="FTPY") %>% 
  ggplot(aes(model, n_events, fill=targets)) +
  geom_boxplot(outlier.size = .1) +
  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
  # ggthemes::scale_fill_solarized() +
  scale_fill_manual(values = palettes_method) +
  theme(legend.position = "none")

plot_rf <- disco_rf %>% 
  filter(ntips>10) %>% 
  # filter(ntips>=length(sptree$tip.label)/2) %>%
  separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
  mutate(model=factor(model, levels=models)) %>% 
  ggplot(aes(model, rf, fill=targets)) + 
  geom_boxplot() + 
  scale_fill_manual(values = palettes_method) +
  labs(x="", y="Norm. RF to species tree", fill="Target sets")

nodes_clades <- fortify(sptree) %>% 
  # filter(!isTip) %>% 
  mutate(ordered = rank(y)) %>% 
  mutate(label_int = fct_reorder(label, y)) %>%
  select(label_int, ordered)

trees_df <- fortify(apro_trees) 
nodes_df <- trees_df %>% 
  group_by(.id) %>% 
  mutate(ordered = rank(y)) %>% 
  left_join(nodes_clades) %>% 
  separate(.id, c("targets", "model")) %>% 
  filter(!isTip, label!="") %>% 
  mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
  separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), 
           ";", convert = TRUE) %>% 
  mutate(freq=f1+f2+f3, 
         model=factor(model, levels=models))

plot_apro <- nodes_df %>%
    filter(model!="FTPY") %>% 
  ggplot(aes(model, q1, fill=targets)) + 
  geom_boxplot() + 
  scale_fill_manual(values = palettes_method) +
  labs(x="", y="Quartet support", fill="Target sets")

# firstrow <- (plot_rf + plot_DL + 
#     plot_layout(guides = "collect"))
# 
# secondrow <- ( + 
#     plot_layout(guides = "collect", widths = c(1,2)))

fig3 <- plot_rf + plot_apro + plot_DL +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"),
        legend.position = "bottom")

ggsave("draft/figures/fig3.pdf", fig3, width = 10, height = 4)
# reco_rnkgs_plot <- reco %>% 
#   group_by(gene, target) %>% 
#   slice_min(n_events, with_ties = T) %>% 
#   group_by(target, model) %>% 
#   count() %>% 
#   ggplot(aes(target, n, fill=model)) + 
#   geom_bar(stat = "identity") +
#   geom_text(aes(label=n), position = position_stack()) +
#   scale_fill_manual(values = palettes_model)
```

```{r FigS1}
plot_bs <- df_bs %>% 
  mutate(targets=target,
         model=factor(model, levels=models)) %>% 
  ggplot(aes(model, support, fill=targets)) +
  geom_boxplot(outlier.size = .1) +
  scale_fill_manual(values = palettes_method) +
  theme_classic() +
  labs(y="Boostrap", x="", fill="") +
  theme(legend.position = "bottom")

var_r2t <- reco %>% 
  filter(model!="FTPY", targets=="common") %>% 
  ggplot(aes(variance_r2t, model)) +
  geom_violin(scale = "width", color="black", fill="grey40") +
  scale_x_continuous(limits = c(0,.5), expand = expansion(mult=0, add = c(0,0.05))) +
  labs(x="Variance Root-to-tip distance") +
  # scale_fill_manual(values = palettes_model) +
  theme(legend.position = 'none',
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,0,0))

rf_plot <- rf_text %>% 
  ggplot(aes(model.x, model.y)) +
  geom_tile(aes( fill=value)) +
  geom_text(aes(label=round(value, 2)), color="white") +
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Median RF") + 
  labs(x="Median RF Distance", y="") +
  theme_classic() +
  theme(legend.position = "none", 
        axis.line = element_blank(),
        axis.ticks = element_blank())

df_red <- select(df_filtered, query, target, singleton)

@@

# for what this should be grouped?
plot_dist <- res %>% 
  filter(query!=target) %>% 
  group_by(query, singleton, model) %>%
  summarise(median_dist=median(nn_dist/max_dist), n=n()) %>% 
  ggplot(aes(median_dist, model,  fill=singleton)) + 
  ggdist::stat_halfeye(alpha=.8, point_colour = NA, .width = 0, height=.6, 
                       position = position_nudge(y=.15)) +
  geom_boxplot(
    width=.2,
    position = position_dodge(),
    # removing outliers
    outlier.color = NA,
    alpha = 1,
    # position= position_nudge(y=-.5)
  ) + 
  labs(x="Normalised internode distance", y="Method", fill="") +
  scale_x_continuous(expand = c(0,0.01), limits = c(0,1)) +
  theme(panel.border = element_rect(fill = NA)) + 
  scale_fill_manual(values = palette_singleton_p) + 
  theme(legend.position = "bottom")

fig_s1 <- ((plot_bs / rf_plot) + 
             plot_layout(heights = c(2,1)) | 
             (plot_dist / var_r2t) + 
             plot_layout(heights = c(2,1))) +
  plot_layout(widths = c(1, 2)) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
ggsave("draft/figures/figS1.pdf", fig_s1, width = 10, height = 6)
```

```{r FigS2}
pp1_plot <- nodes_df %>%
    filter(model!="FTPY") %>% 
    ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
    geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
                   data = . %>% 
                     group_by(model, label_int) %>% 
                     summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
                   position = position_dodge(width=.7), color="grey70") +
    geom_point(aes(shape=targets),
               position = position_dodge(width=.7)) +
  scale_fill_manual(values=palettes_model) +
  scale_color_manual(values = palettes_model_d) +
  scale_shape_manual(values = c(21,22,23,24)) +
  labs(y="Node") +
  guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
         shape = guide_legend(override.aes = list(size=3))) +
  theme(legend.position = "bottom", legend.box="vertical") + 
  theme_bw()

labeled_sptree <- sptree
labeled_sptree$tip.label <- tax$Species[match(labeled_sptree$tip.label, tax$mnemo)]

sptree_plot <- ggtree(labeled_sptree) + 
  geom_tiplab(aes(label=sub("[^A-Z].+? ", ". ", label)),fontface="italic",
              align = TRUE) + 
  geom_nodelab(color="firebrick", 
               nudge_x = -.15, hjust = 1, 
               nudge_y = .3, size=3.5)+
  xlim(c(-4, 21))

fig_s2 <- (sptree_plot + pp1_plot) +
  plot_layout(widths = c(1, 1.5)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
ggsave("draft/figures/figS2.pdf", fig_s2, width = 12, height = 6, device = cairo_pdf)
```
